Ship-based lidar evaluation of Southern Ocean low clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses

### 2025-07-18 manuscript_agu_rev1.tex
Update on Overleaf.

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW) radiation. To first order, the total cloud fraction, cloud phase, and the liquid and ice water path are the most important cloud properties influencing SW and LW radiation. These properties are in turn influenced by the atmospheric thermodynamics, convection and circulation, and both the indirect and direct effects of aerosols. Second-order effects on SW and LW radiation are associated with the cloud droplet size distribution, ice crystal habit, cloud lifetime, and direct radiative interaction with aerosols. In the 6$^\mathrm{th}$ phase of the Coupled Model Intercomparison Project [CMIP6; \citeA{eyring2016}], the cloud feedback has increased relative to CMIP5 \cite{zelinka2020}, especially in the Southern Hemisphere mid-to-high latitudes, which is one of the main reasons for the higher climate sensitivity of CMIP6 models.

The SO is known to be a problematic region for climate model biases \cite{schuddeboom2021,hyder2018,cesana2022,zhao2022} due to a lack of surface and in situ observations. This region has also long been a lower priority region for numerical weather prediction (NWP) and climate model development because of its distance from populated areas. Nevertheless, radiation biases and changes over an area of its size have a substantial influence on the global climate \cite{rintoul2011}, such as affecting the Earth's radiation balance, ocean heat, and carbon uptake \cite{williams2023}, and the SO is also an important part of the global ocean conveyor belt \cite{wang2014b}. In general, marine clouds have a disproportionate effect on top-of-atmosphere (TOA) SW radiation due to the relatively low albedo of the sea surface. The relative longitudinal symmetry of the SO means that model cloud biases tend to be similar across longitudes.

@@

When subsetted by cyclonic and non-cyclonic situations (Fig.~\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a larger amount of observed cloudiness, including peak and total cloud fraction, by about 10\%. In the cyclonic situations, the model vertical profiles of cloud occurrence compare well with observations, but they peak higher by about 200 m and are larger by about 8\%. The reanalyses tend to underestimate cloud occurrence above 1 km by about 5\% and near the surface by about 15\%. Non-cyclonic situations are similar to the general case, also because they form the majority of analyzed profiles (83\%).

>>>>

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW) radiation. To first order, the total cloud fraction, cloud phase, and the liquid and ice water path are the most important cloud properties influencing SW and LW radiation. These properties are in turn influenced by the atmospheric thermodynamics, convection and circulation, and both the indirect and direct effects of aerosols. Second-order effects on SW and LW radiation are associated with the cloud droplet size distribution, ice crystal habit, cloud lifetime, and direct radiative interaction with aerosols \cite{boucher2013}. In the 6$^\mathrm{th}$ phase of the Coupled Model Intercomparison Project [CMIP6; \citeA{eyring2016}], the cloud feedback has increased relative to CMIP5 \cite{zelinka2020}, especially in the Southern Hemisphere mid-to-high latitudes, which is one of the main reasons for the higher climate sensitivity of CMIP6 models.

The SO is known to be a problematic region for climate model biases \cite{schuddeboom2021,hyder2018,cesana2022,zhao2022} due to a lack of surface and in situ observations. This region has also long been a lower priority region for numerical weather prediction (NWP) and climate model development because of its distance from populated areas. Nevertheless, radiation biases and changes over an area of its size have a substantial influence on the global climate \cite{rintoul2011,bodas-salcedo2012}, such as affecting the Earth's radiation balance, ocean heat, and carbon uptake \cite{williams2023}, and the SO is also an important part of the global ocean conveyor belt \cite{wang2014b}. In general, marine clouds have a disproportionate effect on top-of-atmosphere (TOA) SW radiation due to the relatively low albedo of the sea surface. The relative longitudinal symmetry of the SO means that model cloud biases tend to be similar across longitudes.

@@

When the data are subset as either cyclonic and non-cyclonic situations (Fig.~\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a larger amount of observed cloudiness, including peak and total cloud fraction, by about 10\%. In the cyclonic situations, the model vertical profiles of cloud occurrence compare well with observations, but they peak higher by about 200 m and are larger by about 8\%. The reanalyses tend to underestimate cloud occurrence above 1 km by about 5\% and near the surface by about 15\%. Non-cyclonic situations are similar to the general case, also because they form the majority of analyzed profiles (83\%).
